{\boldmath Pseudoscalar top-Higgs coupling: Exploration of $\mathrm{CP}$-odd observables to resolve the sign ambiguity }

### 2016-03-01 Paper_ttbarH_Draft.tex
KK edits to Sec. IIIA.

Before continuing on to our analysis, let us make a few comments
regarding the values that we choose for $\kp$ and $\kpt$.
First of all, we note that if the pseudoscalar
coupling $\kpt$ is the only source of physics beyond the SM,
then indirect contraints (based on Higgs production via gluon fusion,
followed by Higgs decay to two photons {\bf (KK: can we just say
  something like $gg\to H \to \gamma\gamma$?  Or is that not
  correct?)}) disfavour $\kp < 0$ but do not
resolve the degeneracy {\bf (KK: did I reword this correctly?)}
in the sign of $\kpt$ \cite{Guadagnoli}. On
the other hand, if one assumes that the tensor structure of the
Higgs interactions are the
same as those of the SM and if one parameterizes these interactions via
one universal Higgs coupling to vector bosons, $\kappa_V$, and one
universal Higgs coupling to fermions, $\kappa_f$, then the measured signal
strengths provided by the ATLAS and CMS collaborations are compatible with
the values predicted by the SM, (namely, $\kappa_f=1$ and $\kappa_V=1$).
With these facts in mind, we will, for the most part,
set the value of the scalar coupling to its SM
value ($\kp=1$) and will allow the pseudoscalar coupling to take on
various values (including both possible signs). In particular, we
analyze the cases $\kpt=0,\pm 0.25, \pm 0.5, \pm 0.75,\pm 1$.
In addition, we also provide some analysis regarding the pure
$\mathrm{CP}$-odd case ($\kp=0,\kpt=1$).
\subsection{Asymmetry}
%\bigskip
\label{sec3.1}
The first type of $\mathrm{CP}$-odd observable we will consider is the
asymmetry. For a certain TP, we define an asymmetry which gives the
balance between the number of events with $\epsilon > 0$ and those for
which $\epsilon < 0$:
\label{eq21}
\mathcal{A}(\epsilon)=\frac{N(\epsilon > 0)-N(\epsilon < 0)}{N(\epsilon > 0)+N(\epsilon < 0)}.
Based on the general expression given in Eq.~(\ref{eq20}), we expect
the following shape for the asymmetry,
\label{eq22}
\mathcal{A}(\epsilon)=\frac{A\kp\kpt}{B\kappa^2_t+C\tilde{\kappa}^2_t},
which for $\kp=1$ can be parameterized as
\label{eq23}
\mathcal{A}(\epsilon)=\frac{a\kpt}{1+b\tilde{\kappa}^2_t},
where the parameter $a\equiv A/B$ determines the sensitivity to the
pseudoscalar coupling, whereas $b\equiv C/B$ quantifies the deviation
from the linear behaviour.  The results for $\kp=1$ and $\kpt=0,\pm 1$
are shown in Table \ref{table1}, where the values in terms of the
corresponding statistical uncertainty are also included. It can be
seen that the considered asymmetries are useful to separate the
$\mathrm{CP}$-mixed cases from the SM case and also to discriminate
the different sign of $\kpt$, giving deviations of order $10\sigma$ in
the former case and $20\sigma$ in the latter.\par The sensitivity of
the asymmetry is quite similar for the three TPs, as can be seen by
including other values of $\kpt$ and using the expression in
Eq.~(\ref{eq23}) as fitting function (see Fig.~\ref{fig3}). In fact,
we obtain the fit parameters $(a=-0.057\pm 0.006, b=0.5\pm 0.2),
(a=-0.056\pm 0.006, b=0.5 \pm 0.2)$ and $(a=0.058\pm 0.006, b=0.6 \pm
0.2)$ for $\epsilon_1$, $\epsilon_2$ and $\epsilon_3$,
respectively. In addition, we include in Fig.~\ref{fig4} the asymmetry
for $\epsilon_1$ isolating the $gg$-initiated, $q\qbar$-initiated and
$pp$ production. We see that the asymmetry is enhanced for
$gg$-initiated production, while it is reduced and of opposite sign
for the $q\qbar$-initiated case. The asymmetry for the $pp$ case is
then dominated by the $gg$ contribution, albeit smaller due to the
$q\qbar$ contribution.
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
\caption{Asymmetries obtained by using $10^5$ simulated events for the
  TPs $\epsilon_1=\TPa, \epsilon_2=\TPb$ and $\epsilon_3=\TPc$ in the
  SM case and the two $\mathrm{CP}$-mixed cases defined by
  $\kp=1,\kpt=\pm 1$.}
\label{table1}
$\kappa_t$&$\tilde{\kappa}_t$~~&$\mathcal{A}(\epsilon_1)$~~&$\mathcal{A}(\epsilon_1)/\sigma_{\mathcal{A}}$& $\mathcal{A}(\epsilon_2)$~~&$\mathcal{A}(\epsilon_2)/\sigma_{\mathcal{A}}$&$\mathcal{A}(\epsilon_3)$~~&$\mathcal{A}(\epsilon_3)/\sigma_{\mathcal{A}}$  \\ 
$1$ & $-1$~~~ & $0.0315$~ & $10.0$ & $0.0332$~ & $10.5$~ & $-0.0307$~~~ & $-9.7$~~~\\[0.6mm]
$1$ & $0$ & $-0.0021$~~~ & $-0.7$~~~ & $0.0009$~ & $0.3$~ & $-0.0011$~~~ & $-0.3$~~~\\[0.6mm]
$1$ & $1$ & $-0.0379$~~~ & $-12.0$~~~ & $-0.0411$~~~& $-13.0$~~~ & $\,0.0378$~ & $12.0$ \\[0.6mm]
%%%%%%%%%%%%%%%%%%%%%%%%% FIGURA 3 %%%%%%%%%%%%%%%%%%%%%%%%%
\vspace*{2.5mm}
\subfloat{\includegraphics[scale=0.45]{ATP1_nuevo.pdf}}
\hspace*{0.002\textwidth}
\subfloat{\includegraphics[scale=0.45]{ATP2_nuevo.pdf}} \\
\subfloat{\includegraphics[scale=0.45]{ATP3_nuevo.pdf}}
\caption{Asymmetries for the TPs
  $\epsilon_1=\epsilon(t,\tbar,n_t,n_{\tbar})$ (top-left),
  $\epsilon_2=\epsilon(Q,\tbar,n_t,n_{\tbar})$ (top-right) and
  $\epsilon_3=\epsilon(Q,t,n_t,n_{\tbar})$ (bottom). The points
  represent the values for $\kpt=0,\pm 0.25, \pm 0.5, \pm 0.75,\pm 1$
  and the red solid line is the fitting curve.}
\label{fig3}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%% FIGURA 4  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace*{-4mm}
\includegraphics[scale=0.45]{ATP1juntos_nuevo.pdf}
\caption{Asymmetry for the TP
  $\epsilon_1=\epsilon(t,\tbar,n_t,n_{\tbar})$. The dashed line (red)
  corresponds to $gg$-initiated production, the dot-dashed line (grey)
  to $q\qbar$-initiated production and the solid line (blue) to $pp$
  production.}
\label{fig4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace*{-6mm} \par On the other hand, we have also tested various
linear combinations of the TPs $\epsilon_{1,2,3}$, with the result
that the asymmetry is enhanced in the following case:

@@

Note that in the $Q$ rest frame $\epsilon_4=Q^0

@@

$(\vec{t}-\vec{\tbar}\,)\cdot(\vec{n}_t\times \vec{n}_{\tbar})$. The
values obtained for the asymmetry of this TP are shown in
\ref{table2}. We see by comparing the results in Tables \ref{table1}
and \ref{table2} that the capability of the asymmetry to separate the

@@

\par Finally, it is worth noting that the asymmetry is not useful for

@@

pseudoscalar hypothesis ($\kp=0,\kpt=1$) because, being linear in both
$\kp$ and $\kpt$, it is expected to vanish in these cases. However, we
will show in the next section that there exist angular distributions
derived from the TPs that are actually suitable to distinguish these

>>>>

Before continuing on to our analysis, let us make a few comments
regarding the values that we choose for $\kp$ and $\kpt$.
First of all, we note that if the pseudoscalar
coupling $\kpt$ is the only source of physics beyond the SM,
then indirect contraints (based on Higgs production via gluon fusion,
followed by Higgs decay to two photons {\bf (KK: can we just say
  something like $gg\to H \to \gamma\gamma$?  Or is that not
  correct?)}) disfavour $\kp < 0$ but do not
resolve the degeneracy {\bf (KK: did I reword this correctly?)}
in the sign of $\kpt$ \cite{Guadagnoli}. On
the other hand, if one assumes that the tensor structure of the
Higgs interactions are the
same as those of the SM and if one parameterizes these interactions via
one universal Higgs coupling to vector bosons, $\kappa_V$, and one
universal Higgs coupling to fermions, $\kappa_f$, then the measured signal
strengths provided by the ATLAS and CMS collaborations are compatible with
the values predicted by the SM, (namely, $\kappa_f=1$ and $\kappa_V=1$).
With these facts in mind, we will, for the most part,
set the value of the scalar coupling to its SM
value ($\kp=1$) and will allow the pseudoscalar coupling to take on
various values (including both possible signs). In particular, we
analyze the cases $\kpt=0,\pm 0.25, \pm 0.5, \pm 0.75,\pm 1$.
In addition, we also provide some analysis regarding the pure
$\mathrm{CP}$-odd case ($\kp=0,\kpt=1$).
%\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Asymmetry}
%\bigskip
\label{sec3.1}

The first type of $\mathrm{CP}$-odd observable that we will consider is an
asymmetry that compares the number of events
for which a given TP is positive to that for which it is negative.
Normalizing to the total number of events, we define
%
\beq
\label{eq21}
\mathcal{A}(\epsilon)=\frac{N(\epsilon > 0)-N(\epsilon < 0)}{N(\epsilon > 0)+N(\epsilon < 0)}.
\eeq 
%
By construction, $\mathcal{A}\in [-1,+1]$.
Based on the general expression given in Eq.~(\ref{eq20}), we expect
the following functional form for the asymmetry,
%
\beq
\label{eq22}
\mathcal{A}(\epsilon)=\frac{A\kp\kpt}{B\kappa^2_t+C\tilde{\kappa}^2_t},
\eeq 
%
which for $\kp=1$ can be parameterized as
%
\beq
\label{eq23}
\mathcal{A}(\epsilon)=\frac{a\kpt}{1+b\tilde{\kappa}^2_t},
\eeq 
%
where the parameter $a\equiv A/B$ determines the sensitivity to the
pseudoscalar coupling, whereas $b\equiv C/B$ quantifies the deviation
from linear behaviour.

Table~\ref{table1} shows numerical results for the
asymmetries associated with three different TPs,
$\epsilon_1$, $\epsilon_2$ and $\epsilon_3$, taking
$\kp=1$ and $\kpt=0,\pm 1$.  The asymmetry $\mathcal{A}$
is shown in each case, along with $\mathcal{A}/\sigma_{\mathcal{A}}$, where
$\sigma_{\mathcal{A}}$ is the corresponding
statistical uncertainty.  As is
evident from the table, the asymmetries in question provide
a clear separation between the SM and the 
$\mathrm{CP}$-mixed cases, with typical deviations being of
order $10\sigma$.  Furthermore, the asymmetries
for the SM case are each statistically consistent with zero,
as one would expect.
The three asymmetries also allow one to
determine the sign of $\kpt$,
with deviations between the $\kpt = \pm 1$ cases typically
being of order $20\sigma$.
{\bf (KK: I guess what we're calculating here is the
  difference between $\mathcal{A}/\sigma_{\mathcal{A}}$ for
  one scenario and the same thing for another scenario.  But that
  could never actually be measured, right\ldots?  Since the scenario
  is either one or the other\ldots.)}
The sensitivity of
the asymmetry is quite similar for the three TPs, as can be seen by
including other values of $\kpt$ and using the expression in
Eq.~(\ref{eq23}) as a fitting function (see Fig.~\ref{fig3}).
Performing such a fit, 
we obtain $(a=-0.057\pm 0.006, b=0.5\pm 0.2),
(a=-0.056\pm 0.006, b=0.5 \pm 0.2)$ and $(a=0.058\pm 0.006, b=0.6 \pm
0.2)$ for $\epsilon_1$, $\epsilon_2$ and $\epsilon_3$,
respectively.

The results shown in Table~\ref{table1}
and Fig.~\ref{fig3} all assume a $pp$ initial state, which is
actually a combination of events coming from $gg$ and $q\qbar$
initial states.
While this combination of initial states is the appropriate scenario
to consider, it
is interesting to consider the relative contributions to the asymmetry
coming from the
$gg$ and $q\qbar$ initial states.  Figure~\ref{fig4} shows
three curves for the ``$\epsilon_1$'' case, one for $gg$-initiated
events, one for $q\qbar$-initiated events, and one for the usual
combination of these events (the``$pp$'' initial state).
Interestingly, we see from Fig.~\ref{fig4} that the asymmetry for this TP
is enhanced for
$gg$-initiated production, while it is reduced and of opposite sign
for the $q\qbar$-initiated events. The asymmetry for the $pp$ case is
evidently
dominated by the $gg$ contribution, but is somewhat smaller
in magnitude due to the
$q\qbar$ contribution.
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
\renewcommand{\arraystretch}{1.4}
\begin{table}[H]
\caption{Asymmetries obtained by using $10^5$ simulated events for the
  TPs $\epsilon_1=\TPa, \epsilon_2=\TPb$ and $\epsilon_3=\TPc$ in the
  SM case and the two $\mathrm{CP}$-mixed cases defined by
  $\kp=1,\kpt=\pm 1$.}
\label{table1}
\begin{center}
\begin{tabular}{|C{1cm}|C{1cm}||C{2cm}|C{2cm}||C{2cm}|C{2cm}||C{2cm}|C{2cm}|}
%\begin{tabular}{|c|r||r|c||r|c||r|c|}
\hhline{|========|}
%\hhline{|--------|}
$\kappa_t$&$\tilde{\kappa}_t$~~&$\mathcal{A}(\epsilon_1)$~~&$\mathcal{A}(\epsilon_1)/\sigma_{\mathcal{A}}$& $\mathcal{A}(\epsilon_2)$~~&$\mathcal{A}(\epsilon_2)/\sigma_{\mathcal{A}}$&$\mathcal{A}(\epsilon_3)$~~&$\mathcal{A}(\epsilon_3)/\sigma_{\mathcal{A}}$  \\ 
\hhline{|========|} 
%\hhline{|--------|}
$1$ & $-1$~~~ & $0.0315$~ & $10.0$ & $0.0332$~ & $10.5$~ & $-0.0307$~~~ & $-9.7$~~~\\[0.6mm]
\hline
$1$ & $0$ & $-0.0021$~~~ & $-0.7$~~~ & $0.0009$~ & $0.3$~ & $-0.0011$~~~ & $-0.3$~~~\\[0.6mm]
\hline
$1$ & $1$ & $-0.0379$~~~ & $-12.0$~~~ & $-0.0411$~~~& $-13.0$~~~ & $\,0.0378$~ & $12.0$ \\[0.6mm]
\hhline{|========|}
%\hhline{|--------|}
\end{tabular}
\end{center} 
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%% FIGURA 3 %%%%%%%%%%%%%%%%%%%%%%%%%
\begin{center}
\vspace*{2.5mm}
\begin{figure}[H]
%\centering
\hspace*{-0.45cm}
\subfloat{\includegraphics[scale=0.45]{ATP1_nuevo.pdf}}
\hspace*{0.002\textwidth}
\subfloat{\includegraphics[scale=0.45]{ATP2_nuevo.pdf}} \\
%\hspace*{0.03\textwidth}
\centering
\subfloat{\includegraphics[scale=0.45]{ATP3_nuevo.pdf}}
\caption{Asymmetries for the TPs
  $\epsilon_1=\epsilon(t,\tbar,n_t,n_{\tbar})$ (top-left),
  $\epsilon_2=\epsilon(Q,\tbar,n_t,n_{\tbar})$ (top-right) and
  $\epsilon_3=\epsilon(Q,t,n_t,n_{\tbar})$ (bottom). The points
  represent the values for $\kpt=0,\pm 0.25, \pm 0.5, \pm 0.75,\pm 1$
  and the red solid line is the fitting curve.}
\label{fig3}
\end{figure}
\end{center}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%%%%%%%%%%%%%%%%%%%%% FIGURA 4  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{center}
\vspace*{-4mm}
\begin{figure}[H]
\centering
\includegraphics[scale=0.45]{ATP1juntos_nuevo.pdf}
\caption{Asymmetry for the TP
  $\epsilon_1=\epsilon(t,\tbar,n_t,n_{\tbar})$. The dashed line (red)
  corresponds to $gg$-initiated production, the dot-dashed line (grey)
  to $q\qbar$-initiated production and the solid line (blue) to $pp$
  production.}
\label{fig4}
\end{figure}
\end{center}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace*{-6mm}

We have also tested various
linear combinations of the TPs $\epsilon_{1,2,3}$ and have found
that the asymmetry is enhanced for the following combination:

@@

Note that in the $Q$ rest frame, $\epsilon_4=Q^0

@@

$(\vec{t}-\vec{\tbar}\,)\cdot(\vec{n}_t\times \vec{n}_{\tbar})$.
{\bf (KK: Is the result from the previous sentence used somewhere?)}
The
values obtained for the asymmetry associated with this TP are shown in
Table~\ref{table2}. By comparing the results in Tables \ref{table1}
and \ref{table2}, we see that the capability of this
asymmetry to distinguish between the

@@

\par Finally, it is worth noting that the asymmetries
described in this subsection are not useful for

@@

pseudoscalar hypothesis ($\kp=0,\kpt=1$).  Since
the numerators of the {\bf (KK: did I reword this correctly?)}
asymmetries are all linear in both
$\kp$ and $\kpt$, they expected to vanish in these cases. However, we
will show in the next section {\bf (KK: subsection?)}
that there exist angular distributions
derived from the TPs that are actually suitable for distinguishing between these
